AttCRISPR : a spacetime interpretable model for prediction of sgRNA on-target activity

### 2020-11-18 Paper/main.tex
11-18 paper

\begin{figure}[!tpb]%figure3
    \centerline{\includegraphics[width=86mm]{spatialmodule.png}}
    \caption{Details of spatial attention module. A convolution layer is used to generate multi-channel map from $X_{oh}$.
    Then concatenated the output of both max-pooling and average-pooling method and forward it to the last convolution layer.
    A sigmoid function is used to map the final result to a range of zero to one at last, 
    which generates the spatial first-order preference matrix $A_s$.
    }\label{fig:spatialmodule}
\end{figure}

@@

As shown in Figure~\ref{fig:spatialmodule}, given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other words, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, we train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{jinek2012a,cong2013multiplex,mali2013rnaguided}.

\subsubsection{Method in temporal domain}
\begin{figure}[tpb]%figure4
    \centerline{\includegraphics[width=86mm]{temporalmodule.png}}
    \caption{The architecture of temporal domain method in AttCRISPR. 
    The input of the method is embedding sgRNA sequence $X_e$. 
    Then Keys $K$, Values $V$ and Queries $Q$ is generated through a classic encoder-decoder structure which is need for temporal attention module.
    Next, the temporal attention module generates the first-order preference $\tilde{A}$. 
    Each of the row vectors in matrix $\tilde{A}$ represents the base preference of sgRNA at the corresponding position, we use their dot product with the corresponding row vector in embedded $X_e$ to build the score of corresponding position. 
    Hence, a full connection layer is used to weighted average them and achieve the temporal output $y_t$.
    }\label{fig:04}
\end{figure}
\begin{figure}[tpb]
\centerline{\includegraphics[width=86mm]{temporalattentionmodule.png}}
\caption{        
    Details of temporal attention module. 
    To generate the first order preference vector in the $i$ position of sgRNA $\tilde{A}_i$. 
    First, the $i$-th row vector of the querys matrix $Q_i$ is multiplied by the transpose of the keys mateix $K^T$, and apply a softmax function to obtain a preliminary weights vector on the values. 
    Second, to favor the alignment points near $i$, the weights vector obtained is multiplied element-by-element by the $i$-th row vector of the damping matrix $G$, as Equation~(\ref{eq:17}) has shown. 
    $G_{ij}$ can be regarded as the result of place a Gaussian distribution centered around $i$, then sampling the position $j$ (a scaling factor is used to ensure the sum of $G_I$ is $l$). 
    Then we achieve the second-order preference matrix $B$, it is also a weights matrix on the values. So the first-order preference matrix $\tilde{A}$ come from the product of $B$ and values matrix $V$.
    Temporal attention module generates the temporal first-order preference matrix $\tilde{A}$ and the temporal second-order preference matrix $B$.
    }\label{fig:05}
\end{figure}

@@

As Figure~\ref{fig:05} show, in our attention module they are calculated by the following equation
\begin{equation}
\left\{\begin{array}{l}
K_i=Encoder({X_e}_i,\theta_E,K_{i-1})
\\ Q_i=Decoder({X_e}_i,\theta_D,Q_{i-1})
\\V=K 
\end{array}\right.\label{eq:14}
\end{equation}
where vector $K_i$, $Q_i$ denotes the $i$-th row of the matrix $K$ and $Q$ accordingly, $Encoder(\cdot)$ and $Decoder(\cdot)$ are independent GRU units, $\theta_E$ and $\theta_D$ denote all the related parameters of GRU networks accordingly. 
In the actual implementation, we apply the bidirectional GRU networks for better performance, and for the sake of conciseness, we show a conventional GRU network here. 
The function $align(\cdot)$ is as follows
\begin{equation}
B=align(Q,K)\label{eq:16}
\end{equation}
\begin{equation}
B_i=softmax(Q_iK^T)\otimes G_i\label{eq:17}
\end{equation}
\begin{equation}
G_{ij}=\left\{\begin{matrix}
exp(\frac{(i-j)^2}{-2\sigma})&,\left | i-j \right |\leqslant \sigma
\\ 0&,\left | i-j \right |>  \sigma
\end{matrix}\right.\label{eq:18}
\end{equation}
where the matrix $B\in\mathbb{R}^{l\times l}$ is the second-order preference we need, and vector $B_i$ denotes the $i$-th row of the matrix $B$. 
$G\in\mathbb{R}^{l\times l}$ is the damping matrix base on Gaussian function. 
Since a simple belief that the closer the base is to the $i$-th position, the more it affects the $i$-th position normally, we use the damping matrix G to constrain the network learning. 
$\sigma$ represents a threshold of length, any base over this length from the position $i$ is not considered to be affected.
Further, if we think of the values matrix as a vector form of the first-order preference $A$ in Equation~(\ref{eq:04}), we can reach the following equation
\begin{equation}
\tilde{A}=BV\label{eq:19}
\end{equation}
according to above, values matrix $V$ comes from the hidden states of a bidirectional GRU networks, which is usually hard to understand.
While $B$ is the second-order preference matrix obtained by the attention mechanism. 
We believe that the $j$-th dimension of $B_i$, denoted as $B_{ij}$, can reveal the effect of the base at position $j$ on position $i$ in the biological sense.

@@

\subsection{Current prediction methods}

As far as we know, DeepHF, the state-of-the-art method, 
achieved Spearman correlation coefficients of 0.867, 0.862 and 0.860 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively \citep{wang2019optimized}. 
Another tool worth noting is CRISPRpred(SEQ), which is based on conventional machine learning and need complicated human feature engineering exercise \citep{MuhammadRafid2020}. 
They also test their tool with this dataset and achieved Spearman correlation coefficients of 0.838, 0.830 and 0.821 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively (without any hyperparameter tuning).

In order to make the comparison apples to apples, we follow the same strategy as \citeauthor{wang2019optimized} to design our experiments.
To be more specific, each set is shuffled and divided into three parts, 76.5\%, 15\% and 8.5\% of the relevant data was used as the training, test and validation set respectively in a single experiment. 
The experiment is repeated ten times with the results recorded and averaged finally.

@@

Two different experiments are carried out in our work. 
The first one is designed for ablation analysis of AttCRISPR. 
We compare the performance of end2end method (without any hand-crafted biological features) in both spatial and temporal domain. 
Furthermore, we test the ensemble method based on the same strategy to prove that the ensemble method in both spatial and temporal domain can significantly improve the performance.

@@

\section{Results}
\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{baselinewithoutbiofeat.eps}}
    \caption{In the absence of hand-crafted biological features, performance of different algorithms for sgRNA activity prediction. 
    (a)-(c) The performance of Temporal AttCRISPR, Spatial AttCRISPR and Ensemble AttCRISPR. 
    The half-violin plots show the mean and distribution of the Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests. 
    (d)-(f) In the absence of hand-crafted biological features, the performance of all prediction methods in these datasets as far as we know. 
    The $mean \pm s.d.$ of the Spearman correlation coefficient between predicted and measured sgRNA activity scores are shown in the bar plots.}\label{fig:06}
\end{figure}

\begin{figure}[!tpb]%figure4
    \centerline{\includegraphics[width=86mm]{baselinewithbiofeat.eps}}
    \caption{
Performance comparisons for the methods before and after integrating with hand-crafted biological features, 
where DeepHF is the RNN integrated with hand-crafted biological features, 
and StAC is the EnAC integrated with hand-crafted biological features. 
The box plot show the mean and distribution of Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests.
}\label{fig:07}
\end{figure}

@@

In addition, in Table~\ref{Tab:withoutbiofeature}, the performance of other methods without using hand-crafted biological features, are also recorded. 
Regardless of the method we developed, RNN reported by \citeauthor{wang2019optimized}, which can be categorized as the method in temporal domain, 
is the most predictive with Spearman correlation coefficients of 0.856, 0.849, 0.851. 
It's obvious that the ensemble AttCRISPR is better at prediction (Figure~\ref{fig:06} (a-c)). 
Furthermore, the prediction ability of models could be boosted by addition of other hand-crafted biological features, which can't be obtained directly by sequence information.

Further experiment is designed to compare the performance of standard AttCRISPR (hand-crafted biological features are used to improve performance of ensemble AttCRISPR)
 and DeepHF, which is a current state-of-the-art method.
\begin{table}[!tpb]
    \processtable{
    Performance comparisons for different methods in the absence of hand-crafted biological features 
    (take Spearman correlation coefficient as evaluation index)\label{Tab:withoutbiofeature}} 
    {\begin{tabular}{@{}lccc@{}}\toprule
        Method & WT-SpCas9 & eSpCas9(1.1) & SpCas9-HF1\\\midrule
        XGBoost$^*$ & 0.845 & 0.831 & 0.818\\
        MLP$^*$ & 0.842 & 0.846 & 0.844\\
        CNN$^*$ & 0.846 & 0.831 & 0.834\\
        RNN$^*$ & 0.856 & 0.849 & 0.851\\
        TAC & 0.857 & 0.844 & 0.851\\
        SpAC & 0.862 & 0.854 & 0.857\\
        EnAC & \textbf{0.868} & \textbf{0.859} & \textbf{0.862}\\
        CRISPRpred(SEQ)$^{\#}$ & 0.838 & 0.830 & 0.821\\
        \botrule
    \end{tabular}}{}
    \processtable{
    Performance comparisons for the methods before and after integrating with hand-crafted biological features
(take Spearman correlation coefficient as evaluation index)\label{Tab:withbiofeature}} 
    {\begin{tabular}{lcccccc}\toprule
        & \multicolumn{2}{c}{WT-SpCas9} & \multicolumn{2}{c}{eSpCas9(1.1)} & \multicolumn{2}{c}{SpCas9-HF} \\ 
        \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \noalign{\smallskip} 
        Method & Mean  & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} \\ \hline
        RNN$^*$ & 0.856 & 3.33 & 0.849 & 5.00 & 0.851 & 4.11 \\
        EnAC   & 0.868 & 2.66 & 0.859 & 4.66 & 0.862 & 3.19 \\
        DeepHF$^*$ & 0.867 & \textbf{2.37} & 0.862 & 4.24 & 0.860 & 3.21 \\
        StAC   & \textbf{0.872} & 2.55 & \textbf{0.867} & \textbf{3.71} & \textbf{0.867} & \textbf{2.65} \\
     \botrule
\end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: 
The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
In the tables, we use the results reported in the relevant papers as the performance of the method directly.}
\end{table}

@@

@@

As shown in Table~\ref{Tab:withbiofeature}, in the absence of hand-crafted biological features, AttCRISPR has significant advantages over DeepHF in predictability. 
Further, integration with the hand-crafted biological features can also improve the performance of AttCRISPR, and achieve Spearman correlation coefficients of 0.872, 0.867 and 0.867 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively. Meanwhile, DeepHF achieve 0.867, 0.862 and 0.860, respectively. 
After integrating with biological features, the performance gap between AttCRISPR and DeepHF is shortened, while AttCRISPR still has better performance. In addition, we also compare the standard deviation of data obtained in ten tests, which are also shown in Table~\ref{Tab:withbiofeature}. 
It reveal that AttCRISPR is more stable than DeepHF.

In the following sections, we will analyze the insight into activity of sgRNA bring through the attention mechanism at both global and local levels 
to validate the attention module in the AttCRISPR can help us to understand the decisions. 
\subsubsection{Global interpretability}

@@

At local level, we analyze a case (consisting of three sgRNAs as Table~\ref{Tab:optimum} show), 
then we expect AttCRISPR to answer two imortant questions based on the local interpretability. 
First, how can we optimize a sgRNA to have more on-target activity. 
Second, what are the reasons for the low activity of the sgRNA.

For the first question, we input the least-active sgRNA in Table~\ref{Tab:optimum} (with the index of 8493 and the activity of 0.831, call source sgRNA for convenience) into the temporal AttCRISPR. 
The score of each position is obtained based on Equation~(\ref{eq:05}) (the calculated symbol with $W$ is Hadamard product instead of dot product, in order to achieve the result in vector form), 
and the results are shown in Figure~\ref{fig:opt}. 
Acording to Figure~\ref{fig:opt}, scores at position 14 and 16 of source sgRNA are significantly below the base line 
(in fact, the scores at position 6 and 11 is also noteworthy, however, we don't find sgRNA in the dataset for comparison). 
If replaced the T at position 14 with C, would generate the same sgRNA as the one with index of 8491, which is with a activity value of 0.861. 
If replaced the T at position 16 with C, would generate the same sgRNA as the one with index of 8492, which is with a activity value of 0.869. 
Therefore, we could conclude that the local interpretability is helpful for us to optimize the sgRNA without exhaustively search.

The second question we expect AttCRISPR to answer is why it give two low scores at position 14 and 16. 
In practice, we will try to answer this question with second-order preference. 
Let AttCRISPR output the second-order preference matrix $B$ corresponding to the source sgRNA, and show it in Figure~\ref{fig:heatmap}. 
It is obvious that a few unusual bright spots appear in the red box in Figure~\ref{fig:heatmap}, 
which show that the nucleotide at position 15 has a great effect on the score of position 14 and 16 (instead of position 13 or 17, which corresponding position are relatively dim in color). 
As shown in Table~\ref{Tab:optimum}, in source sgRNA there are three consecutive Ts at position 14, 15 and 16, and this may reveal that multiple consecutive Us on sgRNA would leads to low on-target activity of sgRNA, which is consistent with earlier report \citep{wu2014genome-wide}.

@@

\begin{table}[!tpb]
    \processtable{Three sgRNA and their activity \label{Tab:optimum}} 
    {\begin{tabular}{@{}lll@{}}\toprule
        Index & sgRNA & Activity \\\midrule
        8493 & ACATGACTTTGGATTTCCCCAGG & 0.831\\
        8492 & ACATGACTTTGGATTCCCCCAGG & 0.869\\
        8491 & ACATGACTTTGGACTTCCCCAGG & 0.861\\
        \botrule
    \end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: 
    \emph{Activity} of sgRNA represents the activity reported in WT-SpCas9 dataset (Supplementary Data 1). }
\end{table}
\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{local0.eps}}
    \caption{
The scores of sgRNA with index of 8493 obtained by the temporal AttCRISPR. 
The bar plot reveals the score at each position at local level, while the line of dashes reveals the scores at global level. 
At global level we calculated the scores of all sgRNA, then average the scores at each position to obtain the global level which can be used as base line and is helpful to judge the relative height of the scores. 
}\label{fig:opt}
\end{figure}

>>>>

@@

As shown in Supplementary Fig. S1, given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other words, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, we train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{jinek2012a,cong2013multiplex,mali2013rnaguided}.

\subsubsection{Method in temporal domain}
\begin{figure}[tpb]%figure4
    \centerline{\includegraphics[width=86mm]{temporalmodule.png}}
    \caption{The architecture of temporal domain method in AttCRISPR. 
    The input of the method is embedding sgRNA sequence $X_e$. 
    Then Keys $K$, Values $V$ and Queries $Q$ is generated through a classic encoder-decoder structure which is need for temporal attention module.
    Next, the temporal attention module generates the first-order preference $\tilde{A}$. 
    Each of the row vectors in matrix $\tilde{A}$ represents the base preference of sgRNA at the corresponding position, we use their dot product with the corresponding row vector in embedded $X_e$ to build the score of corresponding position. 
    Hence, a full connection layer is used to weighted average them and achieve the temporal output $y_t$.
    }\label{fig:04}
\end{figure}

@@

As Supplementary Fig. S2 show, in our attention module they are calculated by the following equation
\begin{equation}
\left\{\begin{array}{l}
K_i=Encoder({X_e}_i,\theta_E,K_{i-1})
\\ Q_i=Decoder({X_e}_i,\theta_D,Q_{i-1})
\\V=K 
\end{array}\right.\label{eq:14}
\end{equation}
where vector $K_i$, $Q_i$ denotes the $i$-th row of the matrix $K$ and $Q$ accordingly, $Encoder(\cdot)$ and $Decoder(\cdot)$ are independent GRU units, $\theta_E$ and $\theta_D$ denote all the related parameters of GRU networks accordingly. 
In the actual implementation, we apply the bidirectional GRU networks for better performance, and for the sake of conciseness, we show a conventional GRU network here. 
The function $align(\cdot)$ is as follows
\begin{equation}
B=align(Q,K)\label{eq:16}
\end{equation}
\begin{equation}
B_i=softmax(Q_iK^T)\otimes G_i\label{eq:17}
\end{equation}
\begin{equation}
G_{ij}=\left\{\begin{matrix}
exp(\frac{(i-j)^2}{-2\sigma})&,\left | i-j \right |\leqslant \sigma
\\ 0&,\left | i-j \right |>  \sigma
\end{matrix}\right.\label{eq:18}
\end{equation}
where the matrix $B\in\mathbb{R}^{l\times l}$ is the second-order preference we need, and vector $B_i$ denotes the $i$-th row of the matrix $B$. 
$G\in\mathbb{R}^{l\times l}$ is the damping matrix base on Gaussian function. 
Since a simple belief that the closer the base is to the $i$-th position, the more it affects the $i$-th position normally, we use the damping matrix G to constrain the network learning. 
$\sigma$ represents a threshold of length, any base over this length from the position $i$ is not considered to be affected.
Further, if we think of the values matrix as a vector form of the first-order preference $A$ in Equation~(\ref{eq:04}), we can reach the following equation
\begin{equation}
\tilde{A}=BV\label{eq:19}
\end{equation}
according to above, values matrix $V$ comes from the hidden states of a bidirectional GRU networks, which is usually hard to understand.
While $B$ is the second-order preference matrix obtained by the attention mechanism. 
We believe that the $j$-th dimension of $B_i$, denoted as $B_{ij}$, can reveal the effect of the base at position $j$ on position $i$ in the biological sense.

@@

@@

Two different experiments are carried out in our work, which follow the same strategy as \citeauthor{wang2019optimized} to design.
To be more specific, each set is shuffled and divided into three parts, 76.5\%, 15\% and 8.5\% of the relevant data was used as the training, test and validation set respectively in a single experiment. 
The experiment is repeated ten times with the results recorded and averaged finally.

The first one is designed for ablation analysis of AttCRISPR. 
We compare the performance of end2end method (without any hand-crafted biological features) in both spatial and temporal domain. 
Furthermore, we test the ensemble method based on the same strategy to prove that the ensemble method in both spatial and temporal domain can significantly improve the performance.

@@

\begin{table}[!tpb]
    \processtable{
    Performance comparisons for different methods in the absence of hand-crafted biological features 
    (take Spearman correlation coefficient as evaluation index)\label{Tab:withoutbiofeature}} 
    {\begin{tabular}{@{}lccc@{}}\toprule
        Method & WT-SpCas9 & eSpCas9(1.1) & SpCas9-HF1\\\midrule
        XGBoost$^*$ & 0.845 & 0.831 & 0.818\\
        MLP$^*$ & 0.842 & 0.846 & 0.844\\
        CNN$^*$ & 0.846 & 0.831 & 0.834\\
        RNN$^*$ & 0.856 & 0.849 & 0.851\\
        TAC & 0.857 & 0.844 & 0.851\\
        SpAC & 0.862 & 0.854 & 0.857\\
        EnAC & \textbf{0.868} & \textbf{0.859} & \textbf{0.862}\\
        CRISPRpred(SEQ)$^{\#}$ & 0.838 & 0.830 & 0.821\\
        \botrule
    \end{tabular}}{}
    \processtable{
    Performance comparisons for the methods before and after integrating with hand-crafted biological features
(take Spearman correlation coefficient as evaluation index)\label{Tab:withbiofeature}} 
    {\begin{tabular}{lcccccc}\toprule
        & \multicolumn{2}{c}{WT-SpCas9} & \multicolumn{2}{c}{eSpCas9(1.1)} & \multicolumn{2}{c}{SpCas9-HF} \\ 
        \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \noalign{\smallskip} 
        Method & Mean  & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} \\ \hline
        RNN$^*$ & 0.856 & 3.33 & 0.849 & 5.00 & 0.851 & 4.11 \\
        EnAC   & 0.868 & 2.66 & 0.859 & 4.66 & 0.862 & 3.19 \\
        DeepHF$^*$ & 0.867 & \textbf{2.37} & 0.862 & 4.24 & 0.860 & 3.21 \\
        StAC   & \textbf{0.872} & 2.55 & \textbf{0.867} & \textbf{3.71} & \textbf{0.867} & \textbf{2.65} \\
     \botrule
\end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: 
The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
In the tables, we use the results reported in the relevant papers as the performance of the method directly.}
\end{table}

\section{Results}

@@

In addition, in Table~\ref{Tab:withoutbiofeature}, the performance of other methods without using hand-crafted biological features, are also recorded. 
Regardless of the method we developed, RNN reported by \citeauthor{wang2019optimized}, which can be categorized as the method in temporal domain, 
is the most predictive with Spearman correlation coefficients of 0.856, 0.849, 0.851. 
It's obvious that the ensemble AttCRISPR is better at prediction (Supplementary Fig. S3(a-c)). 
Furthermore, the prediction ability of models could be boosted by addition of other hand-crafted biological features, which can't be obtained directly by sequence information.

Further experiment is designed to compare the performance of standard AttCRISPR (hand-crafted biological features are used to improve performance of ensemble AttCRISPR)
 and DeepHF, which is a current state-of-the-art method.

@@

As shown in Table~\ref{Tab:withbiofeature}, in the absence of hand-crafted biological features, AttCRISPR has significant advantages over DeepHF in predictability. 
Further, integration with the hand-crafted biological features can also improve the performance of AttCRISPR, 
and achieve Spearman correlation coefficients of 0.872, 0.867 and 0.867 for WT-SpCas9, eSpCas9(1.1) and SpCas9-HF1, respectively. 
Meanwhile, DeepHF achieve 0.867, 0.862 and 0.860, respectively. 
After integrating with biological features, the performance gap between AttCRISPR and DeepHF is shortened, while AttCRISPR still has better performance (Supplementary Fig. S4). 
In addition, we also compare the standard deviation of data obtained in ten tests, which are also shown in Table~\ref{Tab:withbiofeature}. 
It reveal that AttCRISPR is more stable than DeepHF.

In the following sections, we will analyze the insight into activity of sgRNA bring through the attention mechanism at both global and local levels 
to validate the attention module in the AttCRISPR can help us to understand the decisions.

\subsubsection{Global interpretability}

@@

@@

At local level, we analyze a case (consisting of three sgRNAs as Supplementary Tab. S1 show), 
then we expect AttCRISPR to answer two imortant questions based on the local interpretability. 
First, how can we optimize a sgRNA to have more on-target activity. 
Second, what are the reasons for the low activity of the sgRNA.

For the first question, we input the least-active sgRNA in Supplementary Tab. S1 (with the index of 8493 and the activity of 0.831, call source sgRNA for convenience) into the temporal AttCRISPR. 
The score of each position is obtained based on Equation~(\ref{eq:05}) (the calculated symbol with $W$ is Hadamard product instead of dot product, in order to achieve the result in vector form), 
and the results are shown in Figure~\ref{fig:opt}. 
Acording to Figure~\ref{fig:opt}, scores at position 14 and 16 of source sgRNA are significantly below the base line 
(in fact, the scores at position 6 and 11 is also noteworthy, however, we don't find sgRNA in the dataset for comparison). 
If replaced the T at position 14 with C, would generate the same sgRNA as the one with index of 8491, which is with a activity value of 0.861. 
If replaced the T at position 16 with C, would generate the same sgRNA as the one with index of 8492, which is with a activity value of 0.869. 
Therefore, we could conclude that the local interpretability is helpful for us to optimize the sgRNA without exhaustively search.

The second question we expect AttCRISPR to answer is why it give two low scores at position 14 and 16. 
In practice, we will try to answer this question with second-order preference. 
Let AttCRISPR output the second-order preference matrix $B$ corresponding to the source sgRNA, and show it in Figure~\ref{fig:heatmap}. 
It is obvious that a few unusual bright spots appear in the red box in Figure~\ref{fig:heatmap}, 
which show that the nucleotide at position 15 has a great effect on the score of position 14 and 16 (instead of position 13 or 17, which corresponding position are relatively dim in color). 
As shown in Supplementary Tab. S1, in source sgRNA there are three consecutive Ts at position 14, 15 and 16, and this may reveal that multiple consecutive Us on sgRNA would leads to low on-target activity of sgRNA, which is consistent with earlier report \citep{wu2014genome-wide}.

@@

\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{local0.eps}}
    \caption{
The scores of sgRNA with index of 8493 obtained by the temporal AttCRISPR. 
The bar plot reveals the score at each position at local level, while the line of dashes reveals the scores at global level. 
At global level we calculated the scores of all sgRNA, then average the scores at each position to obtain the global level which can be used as base line and is helpful to judge the relative height of the scores. 
}\label{fig:opt}
\end{figure}
